\section{Public key crypto}

Before Mid-1, we created a CCA-secure scheme.
We assumed that the key $K$ is pre-shared between sender, receiver.

Diffie Hellman key exchange.


Can the key be made public, such that converting from public (encryption) key to the
private (decryption) key is hard?

So, we can publish an encryption key that is public, thereby allowing everyone to
communicate with us.


\subsection{Diffie Hellman SKE(Secret Key Exchange)}
\begin{itemize}
\item There is a group $G$ and a generator $g$ of $G$. Eg: $Z_p^x = <g>$. These are *public*.
\item Alice chooses a random element $a \in G$. Alice sends $g^a$ to Bob.
\item Eve is eavesdropping, all she can see is $g^a$.
\item Bob chooses $b$, sends $g^b$ to Alice.
\item Eve sees $g^b$, cannot find $b$.
\item Key is $g^{ab} = g^a \cdot g^b$
\item key for Eve is $g^{b^a}$ ($g^b$ came from Bob).
\item Key for Bob is $g^{a^b}$ ($g^a$ came from Alice).
\item Now, they both have the key, while an eavesdropper cannot find the key.
\end{itemize}

This is insecure if it is possible to get $g^{ab}$ from $g^a$, $g^b$. Even if discrete log is
hard, there could be some way to use group structure to do this.

This assumption is called 'CDH assumption': given $g^a, g^b$, computing $g^ab$ is hard.


\subsection{Does this satisfy our need? Or, RSA}
This solves key exchange, but not the way we wanted to. We wanted to *publish* the encryption key.

\subsubsection{RSA}
\begin{itemize}
\item $p$ and $q$ are two *large* primes of nearly same length. (today, ~512 bits).
  $n = p \cdot q$. $e \in [1..(p-1)]$, $(e, (p - 1)(q - 1)) = 1$
  $d$ such that $ed = 1 mod (p - 1)(q - 1)$

  Public key: $<N, e>$
  Private key: $<p, q, d>$

\item $Encryption(m) = m^e (mod N)$
\item $Decryption(c) = c^d (mod N)$
\end{itemize}

\subsubsection{Correctness of RSA}
\begin{align}
  dec(enc(m)) &= \\
  c^d(mod N) &= \\
  (m^e)^d (mod N) &= \\
  m^{ed} (mod N)
  = m \text(since ed = (p - 1)(q - 1) = \phi(n))
\end{align}

\paragraph{$phi(N)$}
$phi(N) = pq \text{(all numbers)} - q \text{(multiples of p)} - p \text{(multiples of q)} + 1 \text{(double subtraction of $N$)}$

\paragraph{$a^\phi(n) = 1 (mod n)$}
consider the set $S' = {i_1 a , i_2 a, i_3 a, i_\phi(n) a}$, $S = {i_1, i_2, ..., i_\phi(n)}$.
Show that $S$ and $S'$ are permutations.
QED. (TODO: how to box?)


However, here, we don't publish the key.

\paragraph{RSA assumption}

Given Encrypted message $m^e (mod N)$, and the public key $<N, e>$ we cannot get $m$.


\paragraph{Textbook RSA does not work}
\begin{itemize}
\item RSA is deterministic. Hence, we do not have CPA security.
\item Small key, small N: If Key is small, then, for example, let $m^3 = N$. Now, we can compute cube root $m < \sqrt[3](N)$.
\item Small key:  $c_1 = m^3 mod (N_1)$. $c_2 = m^3 mod (N_2)$. $c_3 = m^3 mod (N_3)$ We are multi-casting
  this message to two people, both of whom have chosen $3$ as their exponent.
  Use chinese remainder theorem. Find $m^3$ in $0 \leq m^3 \leq N_1N_2N_3$.
\end{itemize}

\paragraph{Chinese Remainder Theorem}

Given a family of congruence equations $x = a_i (mod N_i)$, all $N_i, N_j, i \neq j$ are pairwise coprime,
Then we can find $x \in N_1N_2N_3...N_k$.

That is, there is a ring isomorphism:

$Z/N_1Z x Z/N_2Z, Z/N_3Z x ... x Z/N_kZ ~= Z/(N_1N_2N_3 \cdots N_k)Z$

Backwards is obvious, just take modulo $N_1, N_2, N_3, \cdots, N_k$.

Simplest case:

Assume $a_1 = 1$, all other $a_k = 0$. In this case we must set, $x = q.N_2N_3N_4..N_k$.
Let $q = (N_2N_3..N_k)^{-1} (mod N_1)$. Hence, $a_1 = 1 (mod N_1)$. (since $q.N_2N_3N_4..N_k = 1$ (mod $N_1$)).
Also, this number $x$ modulo any *other* $N_k, k \neq 1$ will be $0$ since $x$ is a multiple of that $N_k$.

Similarly, the vector
$[0, 1, 0, 0, ....0]$ = Let  $ Sol= Pi_{i = 1}^k N_i / N_2$. Now, the number we need is $x = Sol*(Sol^{-1}) (mod N_2)$.

So, we can in generate construct our ``basis vectors'' $[1, 0, 0, ...], [0, 1, 0, 0, ..], [0, 0, 1, ...]$.
So, write any number as:
$a_1 [1, 0, 0, ...] + a_2 [0, 1, 0, ...] + a_3 [0, 0, 1, ...]$ (I am somewhat confused, how do we *find* $a_i$?)


\subsection{PKCS v1.5 (Public Key standard)}
We give a probabilistic version of RSA to give it CPA security. To encrypt a message $m$:

\begin{itemize}
\item $Enc(m) = (0000~0000~||~0000 0010~||~r$ (at least 8 bytes, $r \neq all-zeroes$) $||~0000~0000 || m)^e (mod N)$ (|| = concatenation.)
\item We lose *at least* 11 bytes of performance. ($1 + 1 + (\geq 8) + 1$). If we fix length of $r$ to be $8$ (or some other constant),
  then the scheme is insecure! (Homework assignment).
\item for RSA, LSB is the hard core predicate. That is, it is hard to get LSB(x) given $x^e (mod N)$. The first 16 bits are very easy to get (apparently).
  So, we standardise something in the first 16 bits. (WTF? Read the proof of this). The bits after that are right after (where $r$) sits is
  also weaker than LSB. So, we keep the randomness in the weaker bits, and the *actual message* in the stronger bits (remember, LSB is hard!).
\item 
\item Dec(m)
\end{itemize}

Theoretical version of this is called $RSA-OAEP$. (OAEP = optimal asymmetric encryption padding).
This has a proof in the random oracle model that $RSA-OAEP$ is secure. There is no such proof of $PKCS$.

\subsection{El Gamal scheme}
New public key scheme that is based on discrete log, but uses the public key template.
Has a proof of CPA-security in the standard model.


Let $G$ be a group. Let the message be an *element* of the group $m \in G$.
Let $r \in G$, $r$ random. Let the cipher text $c = m \cdot r$. $c \in G$.

We want $r$ to look like $g^{xy}$. We know from diffie-hellman that $g^xy$ cannot be found from
$g^x, g^y$.

Group $G$ is published. Generator $g$ is published ($G = <g>$), $|G|$ is public.
There is a *secret element $x$*. We publish $g^x$.

$PK = <G, g, |G|, h = g^x>$
$SK = x$

$Enc(m) = Choose y \in G. <g^y, h^y * m>$

$Dec(g^y, h^y * m) = (g^x)^y * m = (g^{y^x}) * m$
$m = h^y * m / (g^y)^x.$

So, we can get $m$.

We get $CPA$ security since it is probabilistic (choice of $y$ is probabilistic).


\subsubsection{Homomorphic Encryption with El-Gamal}
Note that El-Gamal is homomorphic WRT group operation.
$<g^y1, h^y1 * m1>, <g^y2, h^y2 * m2>$. Then multiply pointwise. $<g^{y_1 + y_2}, h^{y_1 + y_2}*m_1*m_2>$. Hence, we have encrypted $m_1 * m_2$.




