\chapter{Information Theory, CPA security - Lecture 5}


Most theorems will read as: if X is true, then the protocol $\Pi$ is secure.

\section{Our first example of circumventing an impossibility}

\section{PRNGs - Pseudo random number generators}
This allows us to break $|K| \geq |M|$.
This is still a one-time pad, but it allows us to create $|K| << |M|$.


Deterministic program $G$.  Takes as input $n$-bit string, returns $l(n)$ bit string. We have two
assumptions.
\begin{itemize}
\item 1. $l(n) > n$. Expansion.
\item 2. Pseudorandomess.
\end{itemize}

\subsubsection{Pseudorandomness}
for all PPTM(probabilistic polynomial turing machine) $D$,
$$ |P[D(r) = 1] - P[D(G(s)) = 1]| \leq \negl(|s|) \quad  r \leftarrow \binary^{l(n)} \quad  s \leftarrow \binary^n$$
$r, s$ are chosen uniformly at random. The probability distribution is over the
random coins used by $D$, along with the uniform distributions of $r$ and $s$.


\begin{itemize}
\item Strings of length $l(n)$. pick one at random. probability of picking one of them is
\item Strings of length $n$, and then we inject into $l(n)$ with $G$. Clearly, $|Im(G)| < 2^{l(n)}|$.
So, we can sample all $|Im(G)|$. If we are in a pseudo-random world, it will repeat for sure (with $P = 1$).
If we are in the non-PRNG world (true randomness), the chance that something repeats will be negligibly small.
\item We cannot distinguish with polynomial samples, however. So, PPTM is a good choice for a distinguisher.
\end{itemize}


Given that we have to assume PRNGs exist, there are different ways to proceed:
\begin{itemize}
\item Heuristics - Assume that the PRNG we write is a true PRNG, and then get to work.
\item Specific mathematical assumptions - Assume that certain problems are hard. Build PRNGs from this mathematical assumption.
\item Provable Security - If there exists even one hard problem $P$, then we can use that to build a PRNG.
\item Proven security - prove PRNGs exist.
\end{itemize}


\subsubsection{Assume PRNGs exist. We will build a secure encryption scheme}

% \begin{align*}
%     &M = \binary^{l(m)} \quad K = \binary^m \\
%     &Gen: k = {0, 1}^n$ \\
%     &Enc_k(m) = m \xor G(k) \quad Dec_k(c) = c \xor G(k)
% \end{align*}


Note that this is just one time. If they attacker can see two ciphertexts, they can XOR the ciphertexts to get the XOR of the cleartexts.

\paragraph{Proof that this is sane}.
If the adversary can differentiate between $M_0$ $M_1$, we will use it to break the PRNG (as in, distinguish between PRNG and RNG).
Call the adversary A. It can generate 2 messages $M_0$ and $M_1$. When given $encryption(M_b) = G(k) xor M_b$, he can guess $b = 0 \/ b = 1$ with non-negligible probability.
Call the distinguisher $D$. $D$ has to distinguish between truly random and pseudo random world for our proof.
Given a string $w$ and ask if $w$ can be distinguished by $A$. We can pick $w$ from the PRNG world or the RNG world.

If $A(w xor M0, w xor M1)$ gives us the  correct value(can distinguish), then we are using the PRNG. Otherwise, it is the RNG.


\subsection{Multi message Indistinguishability experiment}

This is defined for an encryption scheme $(\gen, \enc, \dec)$.
\begin{itemize}
\item Adversary outputs a pair of vector of messages $(\vec m_0, \vec m_1)$. Each
    vector contains the same number of messages, and the $i$th messages have the
    same length. That is, $|m_0[i]| = |m_1[i]|$.
\item A random key is created: $ k \leftarrow \gen$, and a random bit
    $b \leftarrow \binary$ is chosen. For all $i$, $c[i] \leftarrow \enc_k(m[i])$
    is computed. $\vec c$ is given to the adversary $A$.
\item The adversary $A$ outputs a bit $b'$. The output of the experiment is $1$
    if $b = b'$, and $0$ otherwise.
\end{itemize}

Security definition of the cryptosystem remains unchanged.


\paragraph{Weakness of one time pads under this threat model}
Note that one time pads will fall to this threat model, since repeatedly
ciphering data with a one-time pad will allow us to extract data from the
one-time pad. Indeed, any deterministic scheme can be attacked under
this threat model. So, we now need probabilistic encryption schemes.

\paragraph{Attacking all deterministic cryptosystems under multi message threat model}
Let $m_0 \equiv (0^n, 0^n), m_1 \equiv (0^n, 1^n)$. Run this through the experiment. We will
be given $c \equiv (c_0, c_1)$. If $c_0 = c_1$, then  we know that the message
was $(0^n, 0^n) = m_0$, and is $m_1$ otherwise. We know this since the encryption
function is deterministic, and  hence $\enc_k(m_0) = \enc_k(m_1) \implies m_0 = m_1$.

\subsection{Secure multiple encryptions using a stream cipher}
TODO: add rigorous definition of stream ciphers.
Two modes of operation.

\paragraph{Synchronized mode} Communicating parties make sure they use
different parts of the stream cipher. This way, no one reuses bits from the
stream cipher. This can be viewed as one giant plaintext, by concatenating
the messages exchanged between the two parties. Security immediately follows.

This is difficult since we need to maintain state between encryptions, to make
sure that we do not reuse randomness.

\paragraph{Unsynchronize mode} We do not need to maintain state, but we
need a more powerful definition of pseudorandom generator. We need a 
seed $s$ and an initialization vector $IV$ of length $n$. The requirement
we will need is that $G(S, IV=iv)$ is pseudo-random (That is, knowing $IV$
while $s$ is secret keeps the output pseudorandom). Furthermore, for two
randomly chosen initialization vectors $IV_1, IV_2$, $G(s, IV_1), G(s, IV_2)$
is pseudo random when $IV_{\{1, 2\}}$ are known, but $s$ is secret. Note that
$s$ is the same for both! We can create an encryption scheme as:

\begin{align*}
    \enc_k(m) \equiv (IV, G(k, IV) \xor m) \qquad
    \dec_k((IV, c)) = m \xor G(k, IV)
\end{align*}

\subsubsection{CPA (Chosen plaintext attack) security: $\privkeysecurity{cpa}(n)$}
This is defined for a cryptosystem $\Pi \equiv (\gen, \enc, \dec)$.
\begin{itemize}
    \item Random key is generated $k \leftarrow \gen(n)$.
    \item The adversary $A$ is given oracle access to $\enc_k(\cdot)$. The
        adversary generates two messages $m_0, m_1$.
    \item Random bit $b \leftarrow \binary$ is chosen. A ciphertext
        $c \leftarrow \enc_k(m_b)$ is computed, and is given to $A$.
    \item $A$ continues to have oracle access to $\enc_k(\cdot)$. Adversary
        outputs a bit $b'$.
    \item Output of experiment is $1$ if $b = b'$, $0$ otherwise.
\end{itemize}

A private key encryption scheme $\Pi \equiv (\gen, \enc, \dec)$ has 
Indistinguishable encryptions under CPA  security  if for all PPTM $A$, there
exists a negligible function $\negl$ such that:

\begin{align*}
    Pr[\privkeysecurity{cpa}(n) = 1] \leq \frac{1}{2} + \negl(n)
\end{align*}


\paragraph{No determistic algorithm can be CPA secure}
The adversary will ask for encryption of $M_0$ and encryption of $M_1$. He gets
back $C_0$ and $C_1$. Then, we can compare that to our result, and find the
random bit $b$.


\paragraph{How to create CPA secure}
$c \equiv \langle r, m \xor \enc_k(r) \rangle$ where $r \leftarrow \binary$ is a random string.
Decryption will never fail. if we know $r$, we can $\xor$ to find $m$. However,
the problem is length doubling: For data of length $n$, we need $|r| = n$.

\subsubsection{Indexable PRNGS}
A PRNG that we can index at a point, and it will start generating from that index. They are called ``pseudorandom functions''.

Consider $Z/pZ^x$. All numbers except $1$ in $Z/pZ$ are generators.

Discrete log: Given $g^x mod p$, given $g$, given $p$, find $x$. (log in a group). We know that Discrete log is hard. Let us try and build a PRNG.


Step 1. Given a PRNG that expands 1 bit, we can use it to create a PRNG that expands any number of bits $n$
$s = seed$. 
$G(s), G(G(s)), G(G(G(s))), G^n(s)$, take the extra bits from each $G^i(s)$. This is a PRNG.

This is a PRNG.

Assume we can break this PRNG. $s_1 s_2 ... s_n$ = stuff from PRNG is distinguishable from $r_1 r_2 r_3 ... r_n$ = Random info.

Construct $s_0 s_1 s_2 .. s_n$, $r_0 s_1 s_2.. s_n$, $r_0 r_1 s_2 s_3 ... s_n$. $r_0 r_1 r_2 r_3 .. r_n$. We know that we can
distinguish first from last. Hence, there must be an adjacent set of strings that can be distinguished, since ``distinguishable'' is transitive (why?)
so, if $r_i  dist r_{i+2}$, we need to have either $r_i dist r_{i + 1}$ or $r_{i + 1} dist r_{i + 2}$. However, between these strings, we have only edited $s_i$.
So, we are able to distinguish one bit extra. This means we can actually distinguish the output of $G$.


Step 2. if we can find $MSB(x)$, we can find x in polynomial time. So, all we need to do is to break $MSB(x)$.


Step 3. Create PRNG that produces one bit output using discrete log.

Take seed s. output $MSB(s_1 = g^s mod p)$. So we now have a PRNG that can create one bit. Second output: $MSB(s_2 = g^{s_1} mod p )$
Third output: $MSB(s_3 = g^{s_3} mod p)$.

Hence, if discrete log is hard, we can get a PRNG. 
