\section{Information Theory, CPA security - Lecture 5}


Most theorems will read as: if X is true, then the protocol $\Pi$ is secure.

\section{Our first example of circumventing an impossibility}

\section{PRNGs - Pseudo random number generators}
This allows us to break $|K| \geq |M|$.
This is still a one-time pad, but it allows us to create $|K| << |M|$.


Deterministic program $G$.  Takes as input $n$-bit string, returns $l(n)$ bit string. We have two
assumptions.
\begin{itemize}
\item 1. $l(n) > n$. Expansion.
\item 2. Pseudorandomess.
\end{itemize}

\subsubsection{Pseudorandomness}
for all PPTM(probabilistic polynomial turing machine) $D$,
$$ |P[D(r) = 1] - P[D(G(s)) = 1]| \leq \negl(|s|) \quad  r \leftarrow \binary^{l(n)} \quad  s \leftarrow \binary^n$$
$r, s$ are chosen uniformly at random. The probability distribution is over the
random coins used by $D$, along with the uniform distributions of $r$ and $s$.


\begin{itemize}
\item Strings of length $l(n)$. pick one at random. probability of picking one of them is
\item Strings of length $n$, and then we inject into $l(n)$ with $G$. Clearly, $|Im(G)| < 2^{l(n)}|$.
So, we can sample all $|Im(G)|$. If we are in a pseudo-random world, it will repeat for sure (with $P = 1$).
If we are in the non-PRNG world (true randomness), the chance that something repeats will be negligibly small.
\item We cannot distinguish with polynomial samples, however. So, PPTM is a good choice for a distinguisher.
\end{itemize}


Given that we have to assume PRNGs exist, there are different ways to proceed:
\begin{itemize}
\item Heuristics - Assume that the PRNG we write is a true PRNG, and then get to work.
\item Specific mathematical assumptions - Assume that certain problems are hard. Build PRNGs from this mathematical assumption.
\item Provable Security - If there exists even one hard problem $P$, then we can use that to build a PRNG.
\item Proven security - prove PRNGs exist.
\end{itemize}


\subsubsection{Assume PRNGs exist. We will build a secure encryption scheme}

% \begin{align*}
%     &M = \binary^{l(m)} \quad K = \binary^m \\
%     &Gen: k = {0, 1}^n$ \\
%     &Enc_k(m) = m \xor G(k) \quad Dec_k(c) = c \xor G(k)
% \end{align*}


Note that this is just one time. If they attacker can see two ciphertexts, they can XOR the ciphertexts to get the XOR of the cleartexts.

\paragraph{Proof that this is sane}.
If the adversary can differentiate between $M_0$ $M_1$, we will use it to break the PRNG (as in, distinguish between PRNG and RNG).
Call the adversary A. It can generate 2 messages $M_0$ and $M_1$. When given $encryption(M_b) = G(k) xor M_b$, he can guess $b = 0 \/ b = 1$ with non-negligible probability.
Call the distinguisher $D$. $D$ has to distinguish between truly random and pseudo random world for our proof.
Given a string $w$ and ask if $w$ can be distinguished by $A$. We can pick $w$ from the PRNG world or the RNG world.

If $A(w xor M0, w xor M1)$ gives us the  correct value(can distinguish), then we are using the PRNG. Otherwise, it is the RNG.


\subsubsection{CPA secure}

Adversary gets to pick $M_0$ and $M_1$, we choose a bit $b$ at random and give encryption of $M_b$. He has an oracle that has oracle access to the encryption algorithm. Even with this, he should not be able to guess $b$.

\paragraph{No determistic algorithm can be CPA secure}

The adversary will ask for encryption of $M_0$ and encryption of $M_1$. He gets back $C_0$ and $C_1$. Then, we can compare that to our result, and find the random bit $b$.


\paragraph{How to create CPA secure}
$C = <R, c xor enc(R)>$ where R is a random string.
Decryption will never fail. if we know R, we can xor twice.
Encryption will not fail because encryption of random data is still random.

We have a problem of length doubling: For one length of data, we need R as well.

\subsubsection{Indexable PRNGS}
A PRNG that we can index at a point, and it will start generating from that index. They are called ``pseudorandom functions''.

Consider $Z/pZ^x$. All numbers except $1$ in $Z/pZ$ are generators.

Discrete log: Given $g^x mod p$, given $g$, given $p$, find $x$. (log in a group). We know that Discrete log is hard. Let us try and build a PRNG.


Step 1. Given a PRNG that expands 1 bit, we can use it to create a PRNG that expands any number of bits $n$
$s = seed$. 
$G(s), G(G(s)), G(G(G(s))), G^n(s)$, take the extra bits from each $G^i(s)$. This is a PRNG.

This is a PRNG.

Assume we can break this PRNG. $s_1 s_2 ... s_n$ = stuff from PRNG is distinguishable from $r_1 r_2 r_3 ... r_n$ = Random info.

Construct $s_0 s_1 s_2 .. s_n$, $r_0 s_1 s_2.. s_n$, $r_0 r_1 s_2 s_3 ... s_n$. $r_0 r_1 r_2 r_3 .. r_n$. We know that we can
distinguish first from last. Hence, there must be an adjacent set of strings that can be distinguished, since ``distinguishable'' is transitive (why?)
so, if $r_i  dist r_{i+2}$, we need to have either $r_i dist r_{i + 1}$ or $r_{i + 1} dist r_{i + 2}$. However, between these strings, we have only edited $s_i$.
So, we are able to distinguish one bit extra. This means we can actually distinguish the output of $G$.


Step 2. if we can find $MSB(x)$, we can find x in polynomial time. So, all we need to do is to break $MSB(x)$.


Step 3. Create PRNG that produces one bit output using discrete log.

Take seed s. output $MSB(s_1 = g^s mod p)$. So we now have a PRNG that can create one bit. Second output: $MSB(s_2 = g^{s_1} mod p )$
Third output: $MSB(s_3 = g^{s_3} mod p)$.

Hence, if discrete log is hard, we can get a PRNG. 

\subsection{Multi message Indistinguishability experiment}

This is defined for an encryption scheme $(\gen, \enc, \dec)$.
\begin{itemize}
\item Adversary outputs a pair of vector of messages $(\vec m_0, \vec m_1)$. Each
    vector contains the same number of messages, and the $i$th messages have the
    same length. That is, $|m_0[i]| = |m_1[i]|$.
\item A random key is created: $ k \leftarrow \gen$, and a random bit
    $b \leftarrow \binary$ is chosen. For all $i$, $c[i] \leftarrow \enc_k(m[i])$
    is computed. $\vec c$ is given to the adversary $A$.
\item The adversary $A$ outputs a bit $b'$. The output of the experiment is $1$
    if $b = b'$, and $0$ otherwise.
\end{itemize}

Security definition of the cryptosystem remains unchanged.


\paragraph{Weakness of one time pads under this threat model}
Note that one time pads will fall to this threat model, since repeatedly
ciphering data with a one-time pad will allow us to extract data from the
one-time pad. Indeed, any deterministic scheme can be attacked under
this threat model. So, we now need probabilistic encryption schemes.

\paragraph{Attacking all deterministic cryptosystems under multi message threat model}
Let $m_0 \equiv (0^n, 0^n), m_1 \equiv (0^n, 1^n)$. Run this through the experiment. We will
be given $c \equiv (c_0, c_1)$. If $c_0 = c_1$, then  we know that the message
was $(0^n, 0^n) = m_0$, and is $m_1$ otherwise. We know this since the encryption
function is deterministic, and  hence $\enc_k(m_0) = \enc_k(m_1) \implies m_0 = m_1$.

