\documentclass[11pt]{book}
%\documentclass[10pt]{llncs}
%\usepackage{llncsdoc}
\usepackage[sc,osf]{mathpazo}   % With old-style figures and real smallcaps.
\linespread{1.025}              % Palatino leads a little more leading
% Euler for math and numbers
\usepackage[euler-digits,small]{eulervm}
\usepackage{physics}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{makeidx}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{listing}
\usepackage{minted}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{mathtools}

\evensidemargin=0.20in
\oddsidemargin=0.20in
\topmargin=0.2in
%\headheight=0.0in
%\headsep=0.0in
%\setlength{\parskip}{0mm}
%\setlength{\parindent}{4mm}
\setlength{\textwidth}{6.4in}
\setlength{\textheight}{8.5in}
%\leftmargin -2in
%\setlength{\rightmargin}{-2in}
%\usepackage{epsf}
%\usepackage{url}

\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{enumitem}
%\usepackage{minted}
%\newminted{fortran}{fontsize=\footnotesize}

\usepackage{xargs}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}

\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=blue,
    filecolor=blue,
    linkcolor=blue,
    urlcolor=blue
}

\usepackage{epsfig}
\usepackage{tabularx}
\usepackage{latexsym}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

\newcommand\ddfrac[2]{\frac{\displaystyle #1}{\displaystyle #2}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\R}{\ensuremath{\mathbb R}}
\newcommand{\coT}{\ensuremath{T^*}}
\newcommand{\Lie}{\ensuremath{\mathfrak{L}}}
\newcommand{\pushforward}[1]{\ensuremath{{#1}_{\star}}}
\newcommand{\pullback}[1]{\ensuremath{{#1}^{\star}}}

\newcommand{\inj}{\hookrightarrow}
\newcommand{\mono}{\inj}


\newcommand{\sur}{\twoheadrightarrow}
\newcommand{\epi}{\sur}

\newcommand{\pushfwd}[1]{\pushforward{#1}}
\newcommand{\pf}[1]{\pushfwd{#1}}

\newcommand{\boldX}{\ensuremath{\mathbf{X}}}
\newcommand{\boldY}{\ensuremath{\mathbf{Y}}}


\newcommand{\G}{\ensuremath{\mathcal{G}}}
\newcommand{\Set}{\ensuremath{\mathbf{Set}} }
% \newcommand{\braket}[2]{\ensuremath{\left\langle #1 \vert #2 \right\rangle}}


\def\qed{$\Box$}
\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{observation}{Observation}
\newtheorem{proof}{Proof}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{exercise}{Exercise}



\title{Topoi, Sheaves, Logic}
\author{Siddharth Bhat}
\date{Monsoon 2019}

\begin{document}
\maketitle
\tableofcontents

\chapter{Topoi of $M$-sets and automata}

Finite automata is just an $M$-set where $M$ is the alphabet.
Subobject classifier for $M$ for a set $A \subseteq X$, $\Xi_A$
maps elements $ x \in X$ 
to the left ideal $\Xi_A(x) \equiv \{ m \in M : m \curvearrowright x \in A \}$.

So, given a "subset automata" $A \subseteq X$, which is a state of states closed under
reachability, $\Xi_A$ assigns to that subset the entire language, and to
other states $x \in X$ the set of strings with prefixes necessary to reach the subset $A$.

What is conjunction and disjunction? Existential and universal quantification?

The adjunction between $Set$ and $Rel$ should allow us to define $M-Rel$, as
well as derive the NFA to DFA algorithm.

\chapter{7 sketches}
I'm following Seven sketches of compositionality. This is a transcription
of the four parts of lecture 7, which promises to cover sheaves, topoi, and
internal logics.

Consider a plane, that has a relationship $\texttt{dial} \mapsto \texttt{thruster}$.
We want this to represent some logical statement 
$\forall t \in Time, @_t (\texttt{dial} = \texttt{bad}), \exists r \in \R, 0 < r < 1, st. @_{t + r} (\texttt{thursters} = \texttt{on})$
We need some category where this logical statement lives. How do we define this
$@$ operator and things like that in the category?

We begin by studying the category of sets, where we normally interpret logic.
Then, we move to richer logics, for example, those with truth values other than
$\top, \bot$. For example, we can have things like $(\top~\text{ for 0 < t < 10})$.

\chapter{Logic in the category Set}

Objects are all sets, arrows are functions.

\section{Properties of \Set}
\begin{itemize}
    \item 1. \Set has limits and colimits. (eg. empty set / initial object, disjoint union / coproduct, pushouts)
    \item 1.1 \Set has terminal objects: the single element set. eg: $1 = \{ \star \}$. (it's unique upto unique isomorphism).
    \item 1.2 \Set has products: $X, Y \in Set \implies X \times Y \in Set$. $X \times Y \equiv \{ (x, y) | x \in X, y \in Y \}$
    \item 1.3 \Set has pullbacks: Given functions $f : X \to A, g: Y \to A$, we can create 
                $X \times_A Y \equiv \{ (x, y) | f(x) = g(y) \}$

        \begin{tikzcd}
            X \times_A Y \arrow[d, dotted, "!"] \arrow[r, dotted, "!"] \arrow[dr, phantom, "\lrcorner", very near start] & X \arrow[d, "f"]\\ 
            Y \arrow[r, "g"] & A
        \end{tikzcd}
        where the $\lrcorner$ means that the square is a pullback square.

    \item 2. Set has epi-mono factorizations: given $f: X \to Y$, we can get
        $epi: X \sur Im(f)$, $mono: Im(f) \inj Y$. $epi$ is surjective, $mono$ is injective.

        \begin{tikzcd}
            X \arrow[rd, "f"] \arrow[r, two heads] & Im(f) \arrow[d, hook]\\ & Y
        \end{tikzcd}
    \item 3. Internal hom: $\Set(A \times B, C) \simeq \Set(A, C^B)$ ($C^B \equiv \text{functions $B \to C$}$),
        where $Set(X, Y)$ is the hom-set.
    \item 4. Subobject classifiers: a \textbf{subobject} of $X \in C$ is (a subset in Set), 
        is an equivalence class of monomorphisms $A \hookrightarrow X$. 
        Given two monomorphisms $f: A \inj X, g: B \inj X$, The subobject will specify if $(f \sim g)$.

        \begin{tikzcd}
            A \arrow[rd, hook] \arrow[d, equal] \\ B \arrow[r, hook] & X
        \end{tikzcd}

        A \textbf{subobject classifier} in $C$ is an object $\Omega$ and a map $1 \rightarrow \Omega$, where $1$ is the terminal object,
        such that for all subobjects, $m: A \mono X$, there exists $\ceil m: X \to \Omega$ such that:

        \begin{tikzcd}
            A \arrow[r, "!"] \arrow[hook, d, "m"]
            \arrow[dr, phantom, "\lrcorner", very near start] & 1 \arrow[d, "true"] \\
            X \arrow[r, "\ceil m"] & \Omega
        \end{tikzcd}

        Where $\ceil m$ is called as the \emph{classifier} of $m: A \rightarrow X$
\end{itemize}

The subobject classifier in $\Set$ is $\Omega = \{ true, false \}$. $(true: 1 \rightarrow \Omega; * \mapsto true)$

We now classify a subobject, the even numbers of \N. let $E = \{ 0, 2 \dots \} \subseteq \N$.
To classify this, we have the commuting square:

\begin{tikzcd}
E \arrow[r, "!"] \arrow[hook, d, "m"]
\arrow[dr, phantom, "\lrcorner", very near start] & 1 \arrow[d, "true"] \\
\N \arrow[r, "\ceil m"] & \Omega
\end{tikzcd}

What is $\ceil m: \N \rightarrow \Omega$? It's going be 
$\ceil e = \begin{cases} true & e \mod 2 = 0 \\ false & \text{otherwise} \end{cases}$.

We need to check that this is indeed a pullback. This clearly commutes, but
we need to check that it's the most general solution.

(It needs to be a pullback so we have a one-to-one correspondence between $E$ and $\ceil e$, apparently.
I don't see it.)

We call a morphism $(X \rightarrow \Omega)$ as a \textbf{predicate}. (This is clear in $\Set$).
Intuitively, previously, $\ceil e$ was a predicate.  The subobject classifier
allows us to find $E$ given the $\ceil e$, thereby find the semantics (as it
were) from the predicate.

\section{Logical operations}

\subsection{And}
\begin{tikzcd}[row sep=large, column sep=15ex]
    1 \arrow[d, hook, "{(true , true)}"]  \arrow[r, equal] \arrow[dr, phantom, "\lrcorner", very near start]  & 1 \arrow[d, "true"]\\
    \Omega \times \Omega \arrow[r, "{ \land \equiv \ceil{(true, true)}}"] & \Omega
\end{tikzcd}

We first draw the LHS, and we get the RHS by apply the subobject classifier
onto the LHS!  $(\land: \Omega \times \Omega \rightarrow \Omega)$. 


\textbf{Question:} How do we compute $\land(false, false)$?

We know that the pullback is ging to be

\chapter{Schloze: Condensed Mathematics Masterclass: Session 1}

YouTube playlist link: \url{https://www.youtube.com/watch?v=PHm4bYziyug&list=PLAMniZX5MiiLXPrD4mpZ-O9oiwhev-5Uq&index=1}


A topos is a generalization of a sheaf. Let $X$ be a topological space. Let $Op(X) \equiv \{ \text{open sets of } X \}$ be a poset 
under inclusion, treated as a category.  A presheaf on $X$ is a functor $F: Op(X)^{op} \rightarrow Set$.
A sheaf on a topological space $X$ is a presheaf $F$ on $X$ such that for any $U \subset X$ open, for $U_i$ open such that
$\cup_i U_i = U$, we have:

\begin{align*}
    &F(U) \xrightarrow{\sim} eq \left(\pi_i F(U_i)^\rightarrow_\rightarrow F(U_i \cap U_j) \right)
\end{align*}

If $Y$ is a topological space, then the presheaf $F_Y(U) \equiv Hom(U, Y)$ is a sheaf.  If $Y$ is a discrete space,
then the sheaf $F_Y$ on $X$ contains the locally constant functions on $X$. This is called as the ``constant sheaf''.
Such a sheaf can also be obtained as the sheafification of the \emph{presheaf} $F_Y(U) \equiv Y$.

\section{Sheafification}
The inclusion $Sh(X) \hookrightarrow Psh(X)$ admits a left adjoint.  This left
adjoint is the sheafification of the presheaf. 

One way to see that this is possible is to notice that the conditions of a sheaf are defined
in terms of limits (products and equalizers). Thus the limit of sheafs is also a sheaf. This ensures
(by the adjoint functor theorem) that this functor has an adjoint.

More concretely, we can construct the sheaf $F^\sharp$ of a presheaf $F$ concretely. Let 

\begin{align*}
&F^\natural(U) \equiv \colim_{(U_i)_i~\text{covers of U}}  eq \left(\prod_i F(U_i)^\rightarrow_\rightarrow \prod_{i, j} F(U_i \cap U_j) \right)
&F^\sharp \equiv (F^\natural)^\natural
\end{align*}

The first operation makes the presheaf into a separated presheaf, and the next time it becomes a sheaf.
\qed.

\section{Stalks}
For any $x \in X$, let $F_x \equiv colim_{x \in U} F(U)$ . These are called as the stalks of $F$ at $x$.
(1) Sheafification does not change stalks. (2) If we have two sections $s, t \in F(U)$, then $s = t$ iff for all $x \in X$, $s_x = t_x$.
So all the information of a sheaf is contained in the stalks. \qed.


\section{Cohomology of abelian sheaves}


Define $AB(X)$ as the sheaf of ableian groups (abelian sheaves) on $X$. This is an abelian category.
We define $0 \rightarrow F' \rightarrow F \rightarrow F'' \rightarrow 0$ is exact iff it is exact at each stalk:
$0 \rightarrow F'_x \rightarrow F_x \rightarrow F''_x \rightarrow 0$.

But globally, we only get $0 \rightarrow F'(x) \rightarrow F(x) \rightarrow F''(X)$.


\section{Sites}

Abstraction of $Op(X)$ with the notion of covers. A Site is a category $C$ together with a set of
maps $Cov(X) \equiv \{ f_i: (X_i \rightarrow X)_i \}$. Properties:

\begin{itemize}
        \item Pullbacks of covers exist and are covers.
        \item Composites of covers are covers.
        \item Isomorphisms are covers.
\end{itemize}

Recall: a presheaf on $C$ is a functor $F: C^{op} \rightarrow Set$. A sheaf on $C$ is a preseaf $F$
such that for all $X \in C$, for all covers, $f_i: X_i \rightarrow X$,  we have that

$$F(X) \simeq eq(\prod_i F(X_i)^\rightarrow_\rightarrow \prod_{i, j}F(X_i \times_X X_j)$$

Notice that we now use fiber products $X_i \times_X X_j$. These fiber products exist
because we asked for pullbacks to exist.  The interesting thing about this definition
is that the morphisms $f_i: X_i \rightarrow X$ don't need to be injections!
Everything carries over, including (a) Sheafification, (b) Abelian sheaves
$Ab(X)$ form an abelian category. 

An example is the opens on $X$. The real first
example is the Etale site of a scheme, where $C \equiv \{ f: Y \rightarrow X \text{ etale} \}$.
covers are maps. Another example is $G$ sets, for $G$ a group, where the covers are actually covers.
Can consider profinite $G$ acting on profinite sets (inverse limit/limit of finite sets), that is the same
as compact totally disconnect Haussdorf spaces.

\section{Morphism of sites}

A morphism of sites $f: C' \rightarrow C$ is a functor $f^{-1}: C \rightarrow C'$ (pullback)
which commutes with finite limits and takes covers to covers. we get a pullback of sheaves
$Sh(C) \rightarrow Sh(C')$.

\section{Grothendieck topos}

A topos is a category $T$ that is equivalent to $Sh(C)$ for some site $C$.
Remark: this admits a characterization in terms of certain axioms on $T$, (due to Giraud).
Sheaves on categories have all limits and colimits, so a topos has all limits and colimits.

A functor of topoi $T' \rightarrow T$ is a pair of adjoint functors $f^*: J \rightarrow J'$,
$f_*: J' \rightarrow J$. $f^*$ preserves all colimns as it is left adjoint. We ask that it also
preserves all finite limits. 

A \emph{point of a topos} is a map of topoi $Set \rightarrow J$. $J$ has enough points if
$\{ f_i^*: J \rightarrow {Set}$ is a conservative family. A functor is conservative if it
reflects isos: That is, $F(f)$ iso implies that $f$ is iso.

\section{Quasicompact}
An object $X \in T$ is quasicompact if for any collection of maps $(f_i: X_i \rightarrow X)$
such that $\cup_i X_i \rightarrow X$ is surjective, then there is some finite $J \subseteq I$
such that $\cup_{j \in J} X_j \rightarrow X$ is surjective. A map $f: Y \rightarrow X$ is quasicompact
if $Y \in T/X$ is quasicompact. (lmao, using slice to relativize!) This is the usual notion of
compact in terms of covers: every cover has a finite subcover.

\section{Quasiseaprated}

$X \in T$ is quasiseparated if for all $Y, Z \rightarrow X$ quasicompact, $Y \times_X Z \rightarrow X$
is quasicompact. Intuitively, we can say that the diagonal is closed. So we may want to say something
like $X$ is quasiseparated if $\Delta: X \rightarrow X \times X$ is quasicompact.




\end{document}
