% // https://tex.stackexchange.com/questions/59702/suggest-a-nice-font-family-for-my-basic-latex-template-text-and-math
\documentclass[5pt]{book}
\usepackage[sc,osf]{mathpazo}   % With old-style figures and real smallcaps.
\linespread{1.025}              % Palatino leads a little more leading

% Euler for math and numbers
\usepackage[euler-digits,small]{eulervm}

%\documentclass[10pt]{llncs}
%\usepackage{llncsdoc}
\usepackage{amsmath,amssymb}
\usepackage{amsthm}
%\usepackage{graphicx}
\usepackage{makeidx}
\usepackage{listing}
\usepackage{comment}
\usepackage{dirtytalk}
\evensidemargin=0.20in
\oddsidemargin=0.20in
\topmargin=0.2in
%\headheight=0.0in
%\headsep=0.0in
%\setlength{\parskip}{0mm}
%\setlength{\parindent}{4mm}
\setlength{\textwidth}{6.4in}
\setlength{\textheight}{8.5in}
%\leftmargin -2in
%\setlength{\rightmargin}{-2in}
%\usepackage{epsf}
%\usepackage{url}



\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{enumitem}
\usepackage{minted}
%\newminted{fortran}{fontsize=\footnotesize}

\usepackage{xargs}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}

\usepackage{hyperref}
\hypersetup{
    colorlinks,
}

\usepackage{epsfig}
\usepackage{tabularx}
\usepackage{latexsym}
\newcommand\ddfrac[2]{\frac{\displaystyle #1}{\displaystyle #2}}

\def\qed{$\Box$}

%\newcommand{\NP}{\texttt{NP}}
%\newcommand{\PSPACE}{\texttt{PSPACE}}
%\newcommand{\NPSPACE}{\texttt{NPSPACE}}
%\newcommand{\TQBF}{\texttt{TQBF}}

\newcommand{\nptime}{\texttt{NP}}
\newcommand{\ptime}{\texttt{P}}
\newcommand{\coptime}{\texttt{co-P}}
\newcommand{\conptime}{\texttt{co-NP}}
\newcommand{\dspace}{\texttt{DPSACE}}
\newcommand{\pspace}{\texttt{PSPACE}}
\newcommand{\logspace}{\texttt{L}}
\newcommand{\nlogspace}{\texttt{NL}}
\newcommand{\conlogspace}{\texttt{co-NL}}
\newcommand{\ph}{\texttt{PH}}
\newcommand{\pathproblem}{\texttt{PATH}}
\newcommand{\copathproblem}{\overline{\texttt{PATH}}}
\newcommand{\clique}{\texttt{CLIQUE}}
\newcommand{\maxclique}{\texttt{MAXCLIQUE}}
\newcommand{\sat}{\texttt{SAT}}
\newcommand{\phtime}{\texttt{PH}}
\newcommand{\ppoly}{$\texttt{P}^{\texttt{poly}}$}
\newcommand{\pathbar}{$\overline{\texttt{PATH}}$}
\newcommand{\problempath}{\texttt{PATH}}
\newcommand{\rp}{\texttt{RP} }
\newcommand{\corp}{\texttt{co-RP} }
\newcommand{\zp}{\texttt{ZP} }
\newcommand{\maxsat}{\texttt{MAXSAT} }
\newcommand{\ilp}{\texttt{ILP} }
\DeclareMathOperator{\shape}{dim}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
%\newtheorem{proof}[theorem]{Proof}

\newcommand{\ip}{\texttt{IP} }
\newcommand{\lp}{\texttt{LP}}
\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\Z}{\ensuremath{\mathbb{Z}}}

\title{Critical thinking}
\author{Siddharth Bhat}
\date{}

\begin{document}

\maketitle
\tableofcontents

\section{Propositions and Arguments}
\begin{definition}
    Logic is the study of the methods and principles used to distinguish 
    correct from incorrect reasoning.
\end{definition}

\begin{definition}
    A proposition asserts that something is the case or it asserts that 
    something is not.
\end{definition}

Questions, commands, and exclamations -- unlike propositions -- are neither 
true nor false.

The term statement is not an exact synonym of proposition, but it is often used
in logic in much the same sense. Some logicians prefer statement to proposition,
although the latter has been more commonly used in the history of logic. Other
logicians eschew both terms as metaphysical, using only the term sentence.

Propositions can be simple and compound.

\begin{itemize}
    \item \textit{Conjunctive proposition}: I am a beatboxer and a computer 
        scientist. (AND)

    \item \textit{Disjunctive proposition}: You can have a scotch or take a 
        hike. (OR)

    \item \textit{Hypothetical proposition}: If God did not exist, it would be 
        necessary to invent him.
\end{itemize}

\begin{definition}
    An argument affirms one proposition, conditioned upon the truth of others.
    The proposition that the argument affirms is called the conclusion and the
    propositions upon which the affirmation of the conclusion depends are called 
    premises, or collectively, the inference.
\end{definition}

Conclusion indicators: therefore, thus, hence, so, accordingly, in consequence,
consequently, proves that, as a result, etc.

Premise indicators: since, because, for, as, follows from, as shown by, etc.

Context is important in understanding arguments. (Duh)

\textbf{Note: Rhetorical questions can serve as premises. Arguments that depend 
on rhetorical questions are always suspect. Because the question is neither true 
nor false, it may be serving as a device to suggest the truth of some proposition 
while avoiding responsibility for asserting it. (Presumption fallacy: plurium 
interrogatium)}

Premises may be considered implicit if they are commonly accepted knowledge.

Explanations look like arguments (same indicators) but are not. (Duh) Figuring 
out which can be iffy.

\begin{definition}
    A deductive argument makes the claim that its conclusion is supported by its
    premises conclusively.
\end{definition}

\begin{definition}
    Inductive arguments are usually based on repeated observation of a phenomenon.
\end{definition}

\begin{definition}
    A deductive argument is said to be valid if the claim made by the premises 
    (if true) provide incontrovertible grounds for the assertion of the 
    conclusion. 
\end{definition}

Examples of validity:

\begin{enumerate}
    \item Valid, true inference and conclusion: All mammals have lungs. All whales 
        are mammals. Therefore all whales have lungs.
    \item Valid, false inference and conclusion: If I owned all the gold in Fort Knox, 
        then I would be wealthy. I do not own all the gold in Fort Knox. Therefore I 
        am not wealthy.
    \item Invalid, true inference, false conclusion: If Bill Gates owned all the gold 
        in Fort Knox, then Bill Gates would be wealthy. Bill Gates does not own all 
        the gold in Fort Knox. Therefore Bill Gates is not wealthy.
    \item Valid, false inference, true conclusion: All fishes are mammals. All whales 
        are fishes. Therefore all whales are mammals.
    \item Invalid, false inference, true conclusion: All mammals have wings. All whales 
        have wings. Therefore all whales are mammals.
    \item Invalid, false inference and conclusion: All mammals have wings. All whales 
        have wings. Therefore all mammals are whales.
\end{enumerate}




\begin{definition}
    Retrograde analysis: Reasoning that seeks to explain how things
    must have developed from things that have happened before.
\end{definition}

\section{Definitions}
\begin{definition}
    Definiendum: The word or symbol being defined
\end{definition}

\begin{definition}
    Definiens: A symbol or group of symbols that have the same definition
    as the Definiendum
\end{definition}

\begin{definition}
    Ambiguity: A term is ambiguous \textbf{in a given context} when it
    has more than one distinct meaning, and the context does not make clear
    what is intended.
\end{definition}

\begin{definition}
    Vagueness: A term is vague, when there are borderline cases where the
    term may or may not apply.
\end{definition}

\subsection{Types of definitions --- (5)}
\texttt{S(tipulative)L(exical)IT(theoretical) P(precising)P(ersuasive)}

\begin{itemize}
    \item \textbf{Stipulative definition}: A new definition where some meaning is
        arbitrarily defined. Such a definition cannot be correct or incorrect.
        Eg. Zetta has been stipulatively defined to mean $10^{12}$.
    \item \textbf{Lexical definition}: Reports the meaning that the Definiendum
        already has. The report maybe correct or incorrect, so such a definition
        maybe true or false.
    \item \textbf{Precising definition}: A definition used to eliminate 
        vagueness or ambiguity.

        It reports on a word that already exists, but it makes the known
        meaning more precise. It could increase the precision by stipulating,
        but the purpose of this stipulation is to \textit{improve a pre-existing}
        meaning.

    \item \textbf{Theoretical definition}: A definition that encapsulates
        an understanding of the theory in which the term is a key element.
    \item \textbf{Persuasive definition}: A definition used to resolve
        disputes  by influencing attitudes and stirring emotion. Often
        uses emotive language.
\end{itemize}

\section{Intention and extension}
\begin{definition} Extension: The set of all objects to which a term
    may be applied. $$extention(P) \equiv \{ x ~|~\forall x \in universe, ~P(x) \}$$
\end{definition}

\begin{definition} Intension: The attributes shared by all and only the objects
    in the class the term denotes. Or, the connotation of the term.

    $$intension(P) \equiv \bigcap_{\forall x \in extension(P)} attribs(x) $$
\end{definition}

\paragraph{The equilateral triangle example} 
Consider equi-\textbf{angular} triangle,
whose intention is different from equi-\textbf{lateral} triangle.

Both of these have the same \textit{extension} (since the sets are the
same), but have different \textit{intentions}.

\section{Extension and denotative definitions}
Means of defining extensive terms:
\begin{definition}Denotative definition: A definition that identifies
    the extention of a term, by (for eg.) listing the members
    of the definition. An extensive definition
\end{definition}


\begin{definition}Ostensive definition: A denotative definition, where
    the definition is made by pointing.

    For example, the word desk means \textit{this}.
\end{definition}

\begin{definition}Quasi Ostensive definition:
    Example: the desk is means \textit{this} article of furniture. This
    presupposes the meaning of \textit{article of furniture}.
\end{definition}

\section{Intention and Intensional Definitions}
\subsection{Types of intentional definitions}
\begin{definition}Subjective intension: Set of all attributes speaker
    believes are posessed by objects denoted by that word
\end{definition}

\begin{definition}Objective intension: Total set of characteristics
    shared by all objects in the word's extension
\end{definition}

\begin{definition}
    Conventional intension: Commonly accepted intension of a term; 
\end{definition}

\subsection{Techniques for creating intensional definitions}
\begin{definition}Synonymous definition: A word, phrase, etc. is defined
    in terms of another
\end{definition}

\begin{definition}Operational definition: Defined by means of operations.
    Example: length can be defined by the measuring procedure.
\end{definition}

\begin{definition}Definitions by genus and difference (Analytical definitions):
    Class whose membership is divided: \textbf{genus}. Subclasses: \textbf{species}.
\end{definition}

\section{Fallacies}
\subsection{Fallacies of relevance - 8} 
\texttt{MISS(ing the point) TWO S(trawman)A(d homenim)F(force)E RED(herring), APPLES(appeal to populace, emotion)}
\begin{itemize}
    \item \textbf{Appeal to populace}: Appeal to people's emotions. Eg. speeches
        by fascist leaders.
    \item \textbf{Appeal to emotion}: Appeal to base emotions such as pity.
    \item \textbf{Red herring}: Distract from the actual argument with a
        \textit{deliberately misleading trail}. Etymology: People who used to
        try and save foxes from being hunted by leaving a smoked herring, which
        confuses dogs, and also turns red.
    \item \textbf{Strawman}: Misconstrue argument to make it seem weaker
        than it actually is, and then defeat the weakened argument.
    \item \textbf{Ad homenim --- abusive type (Argument against the person)}: Attack moral character
        of person. 
    \item \textbf{Ad homenim --- circumstantial type}: Attacking someone's argument
        based on their \textit{circumstance}. Eg. calling a non-vegeterian who
        argues for reduced meat consumption a hypocrite. This does not reduce the
        validity of the argument at all.
    \item \textbf{Appeal to force}: Appeal to threats to coerce the other
        person to accept your argument.
    \item \textbf{Missing the point}: One attacks a different thesis than
        the one the interlocutor was advancing.

\end{itemize}
\subsection{Fallacies of defective induction - (4)}
\texttt{IGNORE AUTHORITY CA(use) GE(generalization)}
\begin{itemize}
    \item \textbf{Argument from ignorance}: Arguing that just because something
        is not \emph{proven} true, it \textit{must be} false, or the converse.
    \item \textbf{Appeal to inappropriate authority}: Appeal to the authority
        of someone who is not an authority on the subject at hand. 

        Example: Invoking Picasso on a discussion about economics.

    \item \textbf{False cause}: Arguing for a cause-and-effect relationship
        where none exists. Eg. you fell sick because of the bees this time
        of year.

    \item \textbf{Hasty Generalization}: Performing induction from a very
        small sample size.
\end{itemize}

\subsection{Fallacies of presumption - (3)}
\texttt{A(ccident) B(egging the question) C(omplex question)}
These come from presuming unjustified assumptions.
\begin{itemize}
    \item \textbf{Accident}: Assuming a generalization applies to all concrete
        cases. 

        Example: It is wrong to steal. We can create corner cases such as 
        \say{what if the person was hungry}? This falls under accident.

    \item \textbf{Complex question}: Constructing a loaded question where
        refuting a part of the question implicitly provides truth to another
        part, which was unintended.

        Example: \say{With all of the hysteria, all of the fear, all of the
        phony science, could it be that man-made global warming is the greatest
    hoax ever perpetrated on the American people?}

    \item \textbf{Begging the question}: Assuming the conclusion in the 
        supposition.

        Example: \say{There is no such thing as knowledge which cannot be
        carried into practice, for such knowledge is really no knowledge at
        all.}
\end{itemize}


\end{document}

